---
title: "RNA-seq in Bioconductor exercises 2"
author: "Rockefeller University, Bioinformatics Resource Centre"
date: "https://rockefelleruniversity.github.io/RU_RNAseq/"
output: 
  html_document:
    number_sections: false  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
params:
  toMessage: false    
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,eval=T)
```

In todays session we will work with some of the RNA-seq data of adult mouse tissues from Bing Ren's lab, Liver and Heart. 
- More information on liver data [can be found here ](https://www.encodeproject.org/experiments/ENCSR000CHB/)
- More information on heart data [can be found here ](https://www.encodeproject.org/experiments/ENCSR000CGZ/)

- Precounted RNAseq reads in genes for these tissues can be found as an R data object in **/Data/geneCounts_Tissue.RData**.

- Salmon quantification for these tissues can be found in directory **/Data/Salmon_Tissue/**.

## RNA-seq analysis.

1. Load the **/Data/geneCounts_Tissue.RData** and create a DESeq2 object. Add 0.25 to every counts and plot a boxplot of the log2 of these updated counts.

```{r, include=FALSE}
library(DESeq2)
library(tximport)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
```

```{r a,cache=TRUE,eval=TRUE,echo=toMessage}
library(DESeq2)
library(tximport)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)

load("data/geneCounts_Tissue.RData")

colData(geneCounts_Tissue)$Tissue <- factor(c("Heart","Heart","Liver","Liver"))

ddsT <- DESeqDataSet(geneCounts_Tissue,design=~Tissue)
boxplot(log2(counts(ddsT,normalized=FALSE)+0.25))
```

2. Run the DEseq workflow function and retrieve normalized counts, add 0.25 to normalised counts and plot a boxplot of the log2 of these updated counts.


```{r b,cache=TRUE,eval=TRUE,dependson="a",warning=FALSE,message=FALSE,echo=toMessage}
ddsT <- DESeq(ddsT)
boxplot(log2(counts(ddsT,normalized=TRUE)+0.25))
```

3. Plot the dispersion fit and estimates  for our DEseq2 object.

```{r c,cache=TRUE,eval=TRUE,dependson="b",warning=FALSE,message=FALSE,echo=toMessage}
plotDispEsts(ddsT)
```

4. Add a new padj value for all genes. Make a barplot of the number of significantly (padj < 0.05) up and down regulated genes by the original padj values and our new padj values..

```{r d,cache=TRUE,eval=TRUE,dependson="b",warning=FALSE,message=FALSE,echo=toMessage}
library(ggplot2)
newRes <- results(ddsT,contrast = c("Tissue","Heart","Liver"))
newResDF <- as.data.frame(newRes)
newResDF$newPadj <- p.adjust(newResDF$pvalue,method = "BH")
newPadj <- ifelse(newResDF$newPadj < 0.05 & newResDF$log2FoldChange > 0,"HeartUp",
              ifelse(newResDF$newPadj < 0.05 & newResDF$log2FoldChange < 0,"HeartDown","NoChange"))
originalPadj <- ifelse(newResDF$padj < 0.05 & newResDF$log2FoldChange > 0,"HeartUp",
              ifelse(newResDF$padj < 0.05 & newResDF$log2FoldChange < 0,"HeartDown","NoChange"))
newPadjFrame <- data.frame(Method="All",Padj=newPadj)
originalPadjFrame <- data.frame(Method="Filtered",Padj=originalPadj)
toPlot <- rbind(newPadjFrame,originalPadjFrame)
ggplot(toPlot,aes(x=Padj,fill=Method))+geom_bar(position="dodge")+theme_bw()
```

5. Produce an MA-plot before and after logFC shrinkage.

```{r e,cache=TRUE,eval=TRUE,dependson="d",warning=FALSE,message=FALSE,echo=toMessage}
newRes <- results(ddsT,contrast = c("Tissue","Liver","Heart"))
DESeq2::plotMA(newRes)
newResLfc <- lfcShrink(dds=ddsT,coef = 2)
DESeq2::plotMA(newResLfc)
```


